---
title: "P8105 Homework 6"
author: "Stephanie Izard"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

##Problem 1

Reading and manipulating dataframe:

* Created variable city_state to include both city and state
* Created binary variable case_status to indicate whether the homicide is solved or unsolved
* Omitted cities Dallas, TX; Phoenix, AZ; and Kansas City, MO because these locations don’t report victim race
* Omitted Tulsa, AL because this is a data entry mistake
* Modified victim_race to have categories white and non-white, with white as the reference category


```{r}
homicide_data <- read_csv("./data/homicide-data.csv", na = c("", "NA", "Unknown")) %>%
  mutate(
    city_state = str_c(city, state, sep = ", "),
    case_status = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest" ~ "unsolved",
      disposition == "Closed by arrest" ~ "solved"),
    victim_race = case_when(
      victim_race == "Asian" ~ "non-white",
      victim_race == "Black" ~ "non-white",
      victim_race == "Hispanic" ~ "non-white",
      victim_race == "Other" ~ "non-white",
      victim_race == "White" ~ "white"),
    victim_race = factor(victim_race, levels = c("white", "non-white"))
  ) %>% 
  filter(city_state != "Dallas, TX", 
         city_state != "Phoenix, AZ", 
         city_state != "Kansas City, MO", 
         city_state != "Tulsa, AL") %>% 
  select(-victim_first, -victim_last)
```

Is the variable victim_age numeric? `r is.numeric(homicide_data$victim_age)`.


