---
title: "P8105 Homework 6"
author: "Stephanie Izard"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
```

##Problem 1

###Reading and manipulating the dataframe:

* Created variable "city_state" to include both city and state
* Created binary variable "resolved" to indicate whether the homicide is solved or unsolved (1 = solved, 0 = unsolved)
* Omitted cities Dallas, TX; Phoenix, AZ; and Kansas City, MO because these locations don’t report victim race
* Omitted Tulsa, AL because this is a data entry mistake
* Modified variable "victim_race" to have categories white and non-white, with white as the reference category
* Made sure the variable "victim_age" was numeric

```{r}
homicide_data <- read_csv("./data/homicide-data.csv", na = c("", "NA", "Unknown")) %>%
  janitor::clean_names() %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_race = case_when(
      victim_race == "Asian" ~ "non-white",
      victim_race == "Black" ~ "non-white",
      victim_race == "Hispanic" ~ "non-white",
      victim_race == "Other" ~ "non-white",
      victim_race == "White" ~ "white")) %>% 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "white")
  ) %>% 
  filter(city_state != "Dallas, TX", 
         city_state != "Phoenix, AZ", 
         city_state != "Kansas City, MO", 
         city_state != "Tulsa, AL") %>% 
  select(-victim_first, -victim_last)
```

###Fitting a logistic regression model for Baltimore, MD:
* Outcome: resolved vs unresolved (case_status)
* Predictors: victim age (victim_age), sex (victim_sex), and race (victim_race)

```{r}
baltimore_glm <- homicide_data %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial())

baltimore_glm %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         ci_lower = exp(estimate - 1.96*std.error),
         ci_upper = exp(estimate + 1.96*std.error)) %>% 
  filter(term == "victim_racenon-white") %>% 
  select(OR, ci_lower, ci_upper) %>% 
  rename("Odds Ratio" = OR,
         "95% CI Lower Bound" = ci_lower,
         "95% CI Upper Bound" = ci_upper) %>% 
  knitr::kable(digits = 3)
```

###Fitting a logistic regression model for all cities:

```{r}
homicide_data_glm <- homicide_data %>% 
  group_by(city_state) %>% 
  nest() %>% 
  mutate(models = map(data, ~glm(resolved ~ victim_age + victim_sex + victim_race, data = .x, family = binomial())),
         models = map(models, broom::tidy)) %>% 
  select(-data) %>% 
  unnest() %>% 
  mutate(OR = exp(estimate),
         ci_lower = exp(estimate - 1.96*std.error),
         ci_upper = exp(estimate + 1.96*std.error)) %>% 
  filter(term == "victim_racenon-white") %>% 
  select(city_state, OR, ci_lower, ci_upper) 

homicide_data_glm %>% 
  rename(City = city_state,
         "Odds Ratio" = OR,
         "95% CI Lower Bound" = ci_lower,
         "95% CI Upper Bound" = ci_upper) %>% 
  knitr::kable(digits = 3)
```

###Plotting estimated ORs and CIs for each city:

* Organized according to descending OR

```{r fig.height = 8, fig.width = 8}
homicide_data_glm %>%
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  group_by(city_state) %>%
  ggplot(aes(x = city_state, y = OR)) +
    geom_point(color = "coral4", size = 1.5) +
    geom_errorbar(aes(ymax = ci_lower, ymin = ci_upper), color = "coral1") +
    labs(x = "City, State",
         y = "Odds Ratio",
         title = "Odds ratios for solving homicides comparing non-white victims to white victims",
         subtitle = "All models are adjusted for victim sex and victim age",
         caption = "Error bars represent the 95% confidence interval") +
    coord_flip() +
    theme_minimal()
```

Comments: Tampa, FL, Birmingham, AL, and Durham, NC have the 3 highest odds ratio for solving homicides comparing non-white victims to white victims, adjusting for victim age and victim sex. These 3 cities are the only cities with an OR estimate greater than 1.0, or having greater odds of a homicide with a non-white victim being solved compared to a homicide with a white victim being solved. Many of the ORs are insignificant becuase their confidence intervals include the null value (1.0).

##Problem 2

####Loading and cleaning the data:

```{r}
bw_df <- read_csv("./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  select(bwt, blength, bhead, malform, gaweeks, babysex) %>% 
  rename(birthweight = bwt,
         length = blength,
         head_circ = bhead,
         gest_age = gaweeks,
         sex = babysex) %>% 
  mutate(malform = as.factor(malform),
         sex = as.factor(sex),
         length = as.numeric(length),
         head_circ = as.numeric(head_circ),
         gest_age = as.numeric(gest_age),
         birthweight = as.numeric(birthweight))

```

The proposed regression model for the factors that underly birthweight is based off of a hypothesized structure, focusing on anthropometric features of the baby. The outcome will be weight at birth in grams (birthweight). Main exposures include baby's length at birth in cm (length), baby's head circumference in cm (head_circ), and the presence of malformations that could affect weight (malform). The model is further adjusted for the variables gestational age in weeks (gest_age), and baby's sex (sex), as these are hypothesized to confound the association between anthropometric features and birthweight.

####Plotting model residuals against fitted values:

```{r}
hyp_model <- lm(birthweight ~ length + head_circ + gest_age + sex + malform, data = bw_df)

bw_df %>% 
  add_predictions(hyp_model) %>% 
  add_residuals(hyp_model) %>% 
  ggplot(aes(x = pred, y = resid)) +
    geom_point() +
    geom_smooth(color = "darkslateblue", se = FALSE) +
    labs(x = "Fitted values", y = "Residuals", title = "Plot of model residuals against fitted values") +
    theme_minimal()
```

Comparing to other models:

```{r fig.height = 7, fig.width = 7}
bw_df_cross <- crossv_mc(bw_df, 100)

bw_df_cross <- bw_df_cross %>% 
  mutate(hyp = map(train, ~lm(birthweight ~ length + head_circ + gest_age + sex + malform, data = .x)),
         comp1 = map(train, ~lm(birthweight ~ length + gest_age, data = .x)),
         comp2 = map(train, ~lm(birthweight ~ length + head_circ + sex + length*head_circ + head_circ*sex + length*sex, data = .x))) %>% 
  mutate(rmse_hyp = map2_dbl(hyp, test, ~rmse(model = .x, data = .y)),
         rmse_comp1 = map2_dbl(comp1, test, ~rmse(model = .x, data = .y)),
         rmse_comp2 = map2_dbl(comp2, test, ~rmse(model = .x, data = .y)))

bw_df_cross %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
    geom_violin(alpha = 0.6) +
    labs(x = "Model", y = "RMSE", title = "Comparing RMSE of hypothesized model to two others") +
    theme_minimal()
    
```

Comments: The hypothesized model (hyp) has the lowest RMSE compared to the other models (comp1 and comp2). Comparative model 1 (using only length at birth and gestational age as predictors) has the highest RMSE, making it the worst option for our data.

